---
title: "Assignment 1"
author: "Akhil Kornala"
date: "2023-09-23"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, comment=NA)
```

1. For this assignment I used the Best Batsman of IPL data, which included both quantitative and qualitative variables. 
Dataset Source: https://www.kaggle.com/datasets/yaranathakur/ipl-all-time-best-batsman

```{r}
library(readr)
Batsman <- read_csv("C:/Users/LENOVO/Desktop/Courses/FML/archive/All-Time-Best-Batsman.csv")
View(Batsman)

```

2. Descriptive Statistics for the variables "Player" and "Runs Scored"
```{r}
summary(Batsman[,c('Player','Runs')])
```

3. Transforming the "Runs" variable into Square Root
```{r}
square_runs <- (Batsman$Runs)^(1/2)
square_runs
```

4. Plotting the variables "Innings Played" and "Runs Scored" against each other
```{r}
plot(Batsman$Inns, Batsman$Runs, main= "Innings Played vs. Runs Scored", xlab='Inns', ylab='Runs')
```

5. Scatter Plot for the variables "Innings Playes" and "Runs Scored"
```{r}
library(ggplot2)
ggplot(data=Batsman, aes(x= Inns,y=Runs))+ geom_point(color = "red", size = 2) +
  labs(title = "Scatter Plot of Innings Played vs. Runs Scored")
```


